### 2023-10-27 manuscript.tex
Update manuscript (#8) [ci skip]
This build is based on
https://github.com/MMV-Lab/im2im-paper/commit/e3d3aa97e37dd21386af3560a76b766120d18211.

This commit was created by the following CI build and job:
https://github.com/MMV-Lab/im2im-paper/commit/e3d3aa97e37dd21386af3560a76b766120d18211/checks
https://github.com/MMV-Lab/im2im-paper/actions/runs/6667059884

This manuscript
(\href{https://MMV-Lab.github.io/im2im-paper/v/ea8d01d4ecd397ce1a10f438734ee13bcba943e9/}{permalink})
was automatically generated
from \href{https://github.com/MMV-Lab/im2im-paper/tree/ea8d01d4ecd397ce1a10f438734ee13bcba943e9}{MMV-Lab/im2im-paper@ea8d01d}
on September 30, 2023.

@@

\begin{itemize}
\item
  \textbf{Justin Sonneck}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-1640-3045}{0000-0002-1640-3045}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/Justin-Sonneck}{Justin-Sonneck}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/JustinSonneck}{JustinSonneck}
  Leibniz-Institut für Analytische Wissenschaften - ISAS - e.V., Dortmund 44139, Germany
\item
  \textbf{Yu Zhou}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0009-0002-3914-1102}{0009-0002-3914-1102}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/audreyeternal}{audreyeternal}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/eternalaudreys}{eternalaudreys}
  Leibniz-Institut für Analytische Wissenschaften - ISAS - e.V., Dortmund 44139, Germany
\item
  \textbf{Jianxu Chen}
  \textsuperscript{\protect\hyperlink{correspondence}{✉}}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-8500-1357}{0000-0002-8500-1357}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/jxchen01}{jxchen01}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/JianxuChen}{JianxuChen}
  Leibniz-Institut für Analytische Wissenschaften - ISAS - e.V., Dortmund 44139, Germany
\end{itemize}

@@

For the 2D problem, we adopted the same network backbone as in the original \emph{EmbedSeg} paper. Example results on a small holdout set of 5 images are shown in Figure \ref{fig:instance}-A (average precision at 50 = 0.866), which is comparable to the original published results {[}\protect\hyperlink{ref-QmYuUQ5K}{11}{]}. For the 3D problem, the original backbone is not directly applicable, due to the aforementioned anisotropic issue and the images in the dataset do not contain enough Z-slices to run through all down sampling blocks in 3D. The anisotropic UNet {[}\protect\hyperlink{ref-jM3v1UjQ}{19}{]} is used here.

Example results were presented in Figure \ref{fig:instance}-B. The green box highlighted a mitotic cell (the DNA signals forming ``spaghetti'' shapes). The average precision at 50 for the fluorescence model is 0.827 and it can be seen that the fluorescence model is able to distinguish the complex DNA signal from the background. Even holes can appear in the predicted segmentation, allowing the prediction of very complex shapes that are theoretically not feasible for other instance segmentation models like StarDist or Cellpose. Additionally, \emph{EmbedSeg}-type models are able to assign spatially unrelated structures to the same instance (see Figure \ref{fig:instance} bottom). Nuclear instance segmentation from brightfield images was much more challenging than from fluorescent images (average precision at 50 = 0.622).

@@

We did a special comparison in this subsection to further illustrate the difference between semantic and instance segmentations. We took the 3D fibrillarin dataset from {[}\protect\hyperlink{ref-5sGcmDuy}{25}{]}. There are multiple channels in each 3D image, including DNA dye, membrane dye, and the structure channel (i.e., fibrillarin). The original fibrillarin segmentation released with the dataset is binary semantic segmentation. With additional cell segmentations available in the dataset, we can know which groups of segmented fibrillarin belong to the same cell. Then, we can convert the semantic segmentation into 3D instance segmentation ground truth (fibrillarin pixels belonging to the same cell are grouped as a unique instance). Sample images and results are shown in Figure \ref{fig:3dseg}. We can observe that the semantic segmentation model is able to achieve good accuracy in determining pixels from the fibrillarin signals (F1 = 0.958 ± 0.008). Meanwhile, the instance segmentation can group them properly (average precision at 50 = 0.795) so that fibrillarin masks from the same cell are successfully identified as unique instances, even without referring to cell segmentation results. This is not a simple distance-based grouping, since the fibrillarin signals from tightly touching nuclei may exist close to each other.

@@

\hypertarget{data-availability}{%
\subsection{Data availability}\label{data-availability}}

@@

>>>>

This manuscript
(\href{https://MMV-Lab.github.io/im2im-paper/v/e3d3aa97e37dd21386af3560a76b766120d18211/}{permalink})
was automatically generated
from \href{https://github.com/MMV-Lab/im2im-paper/tree/e3d3aa97e37dd21386af3560a76b766120d18211}{MMV-Lab/im2im-paper@e3d3aa9}
on October 27, 2023.

@@

\begin{itemize}
\item
  \textbf{Justin Sonneck}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-1640-3045}{0000-0002-1640-3045}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/Justin-Sonneck}{Justin-Sonneck}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/JustinSonneck}{JustinSonneck}
  Leibniz-Institut für Analytische Wissenschaften - ISAS - e.V., Dortmund 44139, Germany; Faculty of Computer Science, Ruhr-University Bochum, Germany
\item
  \textbf{Yu Zhou}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0009-0002-3914-1102}{0009-0002-3914-1102}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/audreyeternal}{audreyeternal}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/eternalaudreys}{eternalaudreys}
  Leibniz-Institut für Analytische Wissenschaften - ISAS - e.V., Dortmund 44139, Germany
\item
  \textbf{Jianxu Chen}
  \textsuperscript{\protect\hyperlink{correspondence}{✉}}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.svg}
  \href{https://orcid.org/0000-0002-8500-1357}{0000-0002-8500-1357}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.svg}
  \href{https://github.com/jxchen01}{jxchen01}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/twitter.svg}
  \href{https://twitter.com/JianxuChen}{JianxuChen}
  Leibniz-Institut für Analytische Wissenschaften - ISAS - e.V., Dortmund 44139, Germany
\end{itemize}

@@

For the 2D problem, we adopted the same network backbone as in the original \emph{EmbedSeg} paper. Example results on a small holdout set of 5 images are shown in Figure \ref{fig:instance}-A (average precision at 50 = 0.866 ± 0.163), which is comparable to the original published results {[}\protect\hyperlink{ref-QmYuUQ5K}{11}{]}. For the 3D problem, the original backbone is not directly applicable, due to the aforementioned anisotropic issue and the images in the dataset do not contain enough Z-slices to run through all down sampling blocks in 3D. The anisotropic UNet {[}\protect\hyperlink{ref-jM3v1UjQ}{19}{]} is used here.

Example results were presented in Figure \ref{fig:instance}-B. The green box highlighted a mitotic cell (the DNA signals forming ``spaghetti'' shapes). The average precision at 50 for the fluorescence model is 0.827 ± 0.082 and it can be seen that the fluorescence model is able to distinguish the complex DNA signal from the background. Even holes can appear in the predicted segmentation, allowing the prediction of very complex shapes that are theoretically not feasible for other instance segmentation models like StarDist or Cellpose. Additionally, \emph{EmbedSeg}-type models are able to assign spatially unrelated structures to the same instance (see Figure \ref{fig:instance} bottom). Nuclear instance segmentation from brightfield images was much more challenging than from fluorescent images (average precision at 50 = 0.622 ± 0.101).

@@

We did a special comparison in this subsection to further illustrate the difference between semantic and instance segmentations. We took the 3D fibrillarin dataset from {[}\protect\hyperlink{ref-5sGcmDuy}{25}{]}. There are multiple channels in each 3D image, including DNA dye, membrane dye, and the structure channel (i.e., fibrillarin). The original fibrillarin segmentation released with the dataset is binary semantic segmentation. With additional cell segmentations available in the dataset, we can know which groups of segmented fibrillarin belong to the same cell. Then, we can convert the semantic segmentation into 3D instance segmentation ground truth (fibrillarin pixels belonging to the same cell are grouped as a unique instance). Sample images and results are shown in Figure \ref{fig:3dseg}. We can observe that the semantic segmentation model is able to achieve good accuracy in determining pixels from the fibrillarin signals (F1 = 0.958 ± 0.008). Meanwhile, the instance segmentation can group them properly (average precision at 50 = 0.795 ± 0.055) so that fibrillarin masks from the same cell are successfully identified as unique instances, even without referring to cell segmentation results. This is not a simple distance-based grouping, since the fibrillarin signals from tightly touching nuclei may exist close to each other.

@@

To enhance the accessibility and traceability of our toolbox, we registered it with https://scicrunch.org/ (RRID: SCR\_024630), https://bio.tools/ (bio.tools ID: biotools:mmv\_im2im) and https://workflowhub.eu/ (https://doi.org/10.48546/workflowhub.workflow.626.1).

\hypertarget{data-and-model-availability}{%
\subsection{Data and model availability}\label{data-and-model-availability}}

@@

\textbf{10. Models and sample data}

To help researchers get started with our tool, we have deposited all models used in the manuscript as well as sample data at https://doi.org/10.5281/zenodo.10034416.
